Automated topic naming to support analysis of software maintenance activities

### 2011-01-13 msr/msr2011.tex
Neil's modifications to the exp2 word lists (and minor other changes).

\author{
\alignauthor 
Abram Hindle\\
   \affaddr{Dept.of Computer Science}\\
   \affaddr{University of California, Davis} \\
   \affaddr{Davis, CA, USA}\\
   \email{abram@softwareprocess.es}
\alignauthor  Neil A. Ernst\\
   \affaddr{Dept. of Computer Science}\\
   \affaddr{University of Toronto}\\
   \affaddr{Toronto, Ontario, CANADA}\\
   \email{nernst@cs.toronto.edu}
\and
\alignauthor Michael W. Godfrey\\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{migod@uwaterloo.ca}
\alignauthor John Mylopoulos \\
   \affaddr{Dept. of Computer Science}\\
   \affaddr{University of Toronto}\\
   \affaddr{Toronto, Ontario, CANADA}\\
   \email{jm@cs.toronto.edu}
}

@@

\newcommand{\shrink}{}

\section{Introduction}

@@

% XXX vanity or space.. 
%XXXXXXXXXXXX OPTIONAL
 Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance techniques. This supports our result that software qualities are not discussed with the same frequency across projects.

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, so the functional requirements would be broadly similar. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains 320 thousand lines of C and C++ source code.  
Choosing an older version of these projects allowed us to focus on projects which have moved into the maintenance phase of the software life-cycle.

@@

Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
While the topic models themselves are generated automatically, how to interpret these topics is less clear.
For example,
% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often
in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
Given the word list \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
The labels are reasonable enough, but still require an expert in the field to determine them. 
We sought to automate the process of naming the topics.

Accordingly, we tried to associate each topic with a label from a list
of keywords and related terms. These keywords were determined before
hand and were not extracted from the project themselves.
We intersected the words of the topics and the words of our word-lists.
We `named' a topic if any of its words matched any of the word-list's words.
A topic could match more than one keyword.
We used several different sets of word-lists for comparison, which we refer to as \textsf{exp1, exp2, exp3} in the text which follows.

@@

Our second word list set, \textsf{exp2}, relied on the ISO quality model, ISO9126~\cite{iso9126}. 
ISO9126 describes six high-level non-functional requirements (listed in Table \ref{tbl:wnsig}). 
There is some debate about the significance and importance of the terms in this model. 
However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
We use these qualities later on as classes in supervised labelling.
The terms extracted from ISO9126 may not capture all words associated with the labels.  
For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. 
We therefore took the words from the taxonomy and expanded them.

\todo[inline]{* clarify the mailinglist stuff we did}
\todo[inline]{This part doesn't make a lot of sense}
To construct these expanded word lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy. 
We then added Boehm's 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 qualities. 
We did the same for the quality model produced by McCall et al. \cite{mccall1977}. 
Finally, we analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.

For the third set of word-lists,  \textsf{exp3}, we extended the word-lists from \textsf{exp2} using unfiltered WordNet similarity matches. 
Similarity in WordNet means siblings in a hypernym tree. 
We do not include these words here for space considerations (but see the Appendix for our data repository). 
It is not clear the words associated with our labels are specific enough. For example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}.

>>>>

\author{
\alignauthor 
Abram Hindle\\
   \affaddr{Dept.of Computer Science}\\
   \affaddr{University of California, Davis} \\
   \affaddr{Davis, CA, USA}\\
   \email{abram@softwareprocess.es}
\alignauthor  Neil A. Ernst\\
   \affaddr{Dept. of Computer Science}\\
   \affaddr{University of Toronto}\\
   \affaddr{Toronto, Ontario, CANADA}\\
   \email{nernst@cs.toronto.edu}
\and
\alignauthor Michael W. Godfrey\\
   \affaddr{David Cheriton School of Computer Science}\\
   \affaddr{University of Waterloo}\\
   \affaddr{Waterloo, Ontario, CANADA}\\
   \email{migod@uwaterloo.ca}
\alignauthor John Mylopoulos \\
   \affaddr{Dept. Information Eng. and Computer Science}\\
   \affaddr{University of Trento}\\
   \affaddr{Trento, Italy}\\
   \email{jm@disi.unitn.it}
}

@@

\newcommand{\shrink}{}
\section{Introduction}

@@

% XXX vanity or space.. 
%XXXXXXXXXXXX OPTIONAL
% NE 0113 this seems too early to mention 'result'.
% Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance techniques. This supports our result that software qualities are not discussed with the same frequency across projects.

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, so the functional requirements would be broadly similar. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains 320 thousand lines of C and C++ source code.  
Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.

@@

Topics are word distributions, lists of words ranked by frequency, that are burdensome to interpret and hard to distinguish and understand. 
While the topic models themselves are generated automatically, how to interpret these topics is less clear.
For example,
% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
in Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
Given the word list \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
The labels are reasonable enough, but still require an expert in the field to determine them. 
Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called `Oracle' in the MySQL case, since Oracle owns MySQL.

In this paper, we had two objectives. One, we sought to automate the process of naming the topics. Two, we sought to label topics with independent terms, in order to facilitate comparison between projects.

Accordingly, we associated each topic with a label from a list
of keywords (`wordlists', in our parlance) and related terms. These wordlists were determined a priori and were not extracted from the projects themselves.
We intersected the words of the topics and the words of our word-lists.
We `labeled' a topic if any of its words matched any of the word-list's words.
A topic could match more than one keyword.
We used several different sets of word-lists for comparison, which we refer to as \textsf{exp1, exp2, exp3} in the text which follows.

@@

Our second word-list, \textsf{exp2}, is derived from the ISO quality model, ISO9126~\cite{iso9126}. 
ISO9126 describes six high-level non-functional requirements (NFRs), maintainability, functionality, portability, efficiency, usability, and reliability. 
There is some debate about the terms in this model. 
However, ISO9126 is ``an international standard and thus provides an internationally accepted terminology for software quality \cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this research.  
%We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
The terms from ISO9126 may not capture all words occuring in the topics that are nonetheless associated with one of the NFRs. 
For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. 
We therefore took the NFRs from the ISO9126 and added to them.

%\todo[inline]{* clarify the mailinglist stuff we did}
%\todo[inline]{This part doesn't make a lot of sense}
To construct these expanded word-lists, we used WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy. 
We then added Boehm's software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven `ilities' into their respective ISO9126 NFRs. 
We did the same for the quality model produced by McCall et al. \cite{mccall1977}. 
We then did a simple random analysis of mailing list messages from an open source project, KDE. If we judged a given message to contain terms that were related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
% Less info is probably better here...
%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.  Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.

For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using unfiltered WordNet similarity matches. 
Similarity in WordNet means siblings in a hypernym tree. 
We do not include these words here for space considerations (but see the Appendix for our data repository). 
It is not clear the words associated with our labels in \textsf{exp3} are specific enough. For example, the label \emph{maintainability} is associated with words \emph{ease} and \emph{ownership}. In general, as we proceed from word-list in \textsf{exp1} to that in \textsf{exp3}, our lists get more generic.
